---
title: "R Notebook"
output: html_notebook
---

# UK Productivity Factors

In this document UK productivity over time will be investigated as a function of UK government spending. The ultimate goal being to answer the 3 following questions:

• Does government spending on factors such as mental health, education, and research and development affect productivity in the UK?
• Is there a relationship between government investment and productivity?
• Can you predict productivity based on investment?

A series of data was provided in order to aid in the investigation. Any justifications or assumptions will be stated when necessary.


Load libriaries
```{r}
library(tidyverse)
library(grid)
```

Load Data
```{r}
uk_productivity_factors <- read_csv("../Deloitte_Investment_Daniel_McGuire/clean_data/uk_productivity_factors.csv") %>%
  mutate(time = as.character(time)) 

uk_productivity_regional <- read_csv("../Deloitte_Investment_Daniel_McGuire/clean_data/uk_regional_productivity.csv")

uk_gdp <- read_csv("clean_data/uk_yearly_gdp.csv")

CC_data <- read_csv("../Deloitte_Investment_Daniel_McGuire/clean_data/CC_data.csv")

OECD_data <- read_csv("../Deloitte_Investment_Daniel_McGuire/clean_data/OECD_data.csv")

# join all international data
all_international_data <- OECD_data %>%
  full_join(CC_data, by = c("location", "time")) %>%
  mutate(time = as.character(time))

# subset for productivity for the whole economy and GDP
uk_productivity_whole <- uk_productivity_regional %>%
  filter(uk_productivity_measure == "UK Whole Economy: Output per hour worked SA: Index 2016 = 100") %>%
  mutate(time = as.character(time)) %>%
  drop_na()

# subset GDP data for seasonally adjusted chained volume measures
uk_gdp_focus <- uk_gdp %>%
  select(c("time", "Gross Domestic Product: chained volume measures: Seasonally adjusted £m")) %>%
  mutate(time = as.character(time)) %>%
  rename("GDP" = "Gross Domestic Product: chained volume measures: Seasonally adjusted £m")


```

The relationship between government spending and individual factors will be investigated first

### Mental Health

```{r}
mental_health_abscences_1 <- uk_productivity_factors %>%
  select("time", "reason_given", "abscences_millions") %>%
  filter(reason_given == "Serious mental health problems" | reason_given == "Stress, depression, anxiety") %>%
  group_by(time, reason_given) %>%
  summarise(abscences = mean(abscences_millions)) %>%
  mutate(time = as.character(time)) %>%
  ggplot() +
  aes(x = time, y = abscences, fill = reason_given) +
  geom_col() +   
  scale_fill_manual(values = c("#fc8d59", "#91bfdb")) + 
  labs(x = "\nYear",
         y = "Abscences Related to Mental Health (millions)",
         title = "Yearly Abscences Related to Mental Health\n",
         fill = "Reason Given") +
  
    theme_classic(base_size = 15) +
    theme(plot.title = element_text(hjust = 0.5), legend.position="bottom")

ggsave("presentation/mental_health_abscences_1.png", plot = mental_health_abscences_1, width = 9)
```


```{r}
all_international_data %>%
  select(c("time", "location", "new_mental_health_units")) %>%
  filter(!is.na(new_mental_health_units)) %>%
  group_by(location) %>%
  summarise("mental_health_units_in_hospitals" = sum(new_mental_health_units)) %>%
  ggplot() + 
  aes(x = reorder(location, -mental_health_units_in_hospitals), y = mental_health_units_in_hospitals) +
  geom_col() +
  coord_flip()
```
```{r}
all_international_data %>%
  filter(time <  2011 & time > 2003,
         labour_productivity_measure == "USD") %>%
  group_by(location) %>%
  summarise(average_labour_productivity = mean(labour_productivity_value)) %>%
  ggplot() + 
    aes(x = reorder(location, -average_labour_productivity), y = average_labour_productivity) +
    geom_col() +
    coord_flip()
```


## Education 

```{r}
further_education_vs_productivity_uk_graph <- all_international_data %>%
  full_join(uk_productivity_whole, by = c("time")) %>%
  select(c("location", "time", "adult_education_level_value", "uk_productivity_value")) %>%
  filter(!is.na(adult_education_level_value), location == "GBR", time < 2017) %>% # year 2017 had a strange value so was omitted
  ggplot() +
  aes(x = adult_education_level_value, y = uk_productivity_value) +
  geom_point(colour = "#91bfdb", size = 2) + 
  geom_smooth(method = "lm", formula = y ~ poly(x, 2), se = FALSE, colour = "#fc8d59", size = 2) + 
  labs(x = "Tertiary Education Rate (% of 25-64 year-old)",
       y = "UK Productivity - Output per Hour Worked",
       title = "Further Education vs Productivity for the UK",
      colour = "Year") +
  theme_classic(base_size = 15) +
  theme(plot.title = element_text(hjust = 0.5))


ggsave("presentation/further_education_vs_productivity_uk.png", plot = further_education_vs_productivity_uk_graph, width = 9)

```

```{r}
education_graph_1 <- all_international_data %>%
  full_join(uk_productivity_whole, by = c("time")) %>%
  full_join(uk_gdp_focus, by = c("time")) %>%
  select(c("location", "time", "tertiary_education_subject", "tertiary_education_exp_value", "uk_productivity_value", "GDP")) %>%
  filter(!is.na(tertiary_education_exp_value), location == "GBR", tertiary_education_subject == "PRY_NTRY", time > 2005) %>% # filter by education investment from primary to post secondary but non tertiary
  mutate(education_spend = (GDP*tertiary_education_exp_value)/(100*1000)) %>% # calculate how much is being spent on education in pounds
  ggplot() +
  aes(y = education_spend, x = time) +
  geom_line(group = 1, colour =   "#91bfdb", size = 2) +
  geom_point(colour =   "#91bfdb", size = 3) +
  labs(y = "\nPrimary-Secondary Education \nInvestment (£ billion)\n",
       title = "Education Investment and Productivity for the UK over Time\n") +
  theme_classic(base_size = 8) +
  theme(axis.title.x=element_blank(), plot.title = element_text(hjust = 0.5))
```


```{r}
education_graph_2 <- all_international_data %>%
  full_join(uk_productivity_whole, by = c("time")) %>%
  full_join(uk_gdp_focus, by = c("time")) %>%
  select(c("location", "time", "tertiary_education_subject", "tertiary_education_exp_value", "uk_productivity_value", "GDP")) %>%
  filter(!is.na(tertiary_education_exp_value), location == "GBR", tertiary_education_subject == "PRY_NTRY", time > 2005) %>% # filter by education investment from primary to post secondary but non tertiary
  # filtered from the year 2005 as the last point was 2000 and the 5 year gap didn't make the analysis clear
  ggplot() +
  aes(x = time, y = uk_productivity_value) +
  geom_line(group = 1, colour =   "#91bfdb", size = 2) +
  geom_point(colour = "#91bfdb", size = 3) +
  labs(x = "Year",
       y = "Output per Hour Worked") +
  theme_classic(base_size = 8) +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
# graph was saved manually
grid.newpage() 
grid.draw(rbind(ggplotGrob(education_graph_1), ggplotGrob(education_graph_2), size = "last"))
```


```{r}
education_investment_vs_productivity_graph <- all_international_data %>%
  full_join(uk_productivity_whole, by = c("time")) %>%
  full_join(uk_gdp_focus, by = c("time")) %>%
  select(c("location", "time", "tertiary_education_subject", "tertiary_education_exp_value", "uk_productivity_value", "GDP")) %>%
  filter(!is.na(tertiary_education_exp_value), location == "GBR", tertiary_education_subject == "PRY_NTRY") %>% # filter by education investment from primary to post secondary but non tertiary
  mutate(education_spend = (GDP*tertiary_education_exp_value)/(100*1000)) %>% # calculate how much is being spent on education in pounds
  distinct() %>%
  ggplot() +
  aes(x = education_spend, y = uk_productivity_value) +
  geom_point(colour = "#91bfdb", size = 3) + 
  geom_smooth(method = "lm", se = FALSE, colour = "#fc8d59", size = 3) + 
   labs(x = "\nPrimary to Secondary Education Investment (£ billions)",
       y = "Output per Hour Worked",
       title = "Education Investment vs Productivity for the UK\n",
      colour = "Year") +
  theme_classic(base_size = 14) +
  theme(plot.title = element_text(hjust = 0.5))

ggsave("presentation/education_investment_vs_productivity_graph.png", plot = education_investment_vs_productivity_graph, width = 11)
```

Repeat with the OECD Data for Labour Productivity
```{r}
all_international_data %>%
  full_join(uk_productivity_whole, by = c("time")) %>%
  full_join(uk_gdp_focus, by = c("time")) %>%
   select(c("location", "time", "tertiary_education_subject", "tertiary_education_exp_value", "labour_productivity_value","labour_productivity_measure", "GDP")) %>%
  filter(!is.na(tertiary_education_exp_value), location == "GBR", tertiary_education_subject == "PRY_NTRY", labour_productivity_measure !="USD", time > 2005) %>% # filter by education investment from primary to post secondary but non tertiary
  mutate(education_spend = (GDP*tertiary_education_exp_value)/(100*1000)) %>% # calculate how much is being spent on education in pounds
  distinct() %>%
  ggplot() +
  aes(x = education_spend, y = labour_productivity_value) +
  geom_point(colour = "#91bfdb", size = 2) + 
  geom_smooth(method = "lm", se = FALSE, colour = "#fc8d59", size = 2) +
  labs(x = "\nPrimary to Secondary Education Investment (£ billions)",
           y = "Labour Productivity (GDP per hour worked)",
           title = "\nPrimary to Secondary Education Investment vs Labour Productivity",
       colour = "Country") +
      theme_classic(base_size = 10) +
      theme(plot.title = element_text(hjust = 0.5))
```


# Research and Development 

```{r}
r_d_graph_1 <- uk_productivity_factors %>%
  full_join(uk_gdp_focus, by = c("time")) %>%
  full_join(uk_productivity_whole, by = c("time")) %>%
  select(c("time", "r_d_exp_measure", "r_d_exp_value","uk_productivity_value", "GDP")) %>%
  filter(r_d_exp_measure == "As a % of GDP") %>%
  mutate(r_and_d_spend = (GDP*r_d_exp_value)/(100*1000)) %>%
  distinct() %>%
  ggplot() +
  aes(x = time, y = r_and_d_spend) +
  geom_line(group = 1, colour =   "#91bfdb", size = 2) +
  geom_point(colour =   "#91bfdb", size = 3) +
  labs(title = "R&D Investment and Productivity over Time",
       y = "R&D Investment (£ billions)") +
    theme_classic(base_size = 10) +
    theme(axis.title.x=element_blank(),plot.title = element_text(hjust = 0.5))
```

```{r}
r_d_graph_2 <- uk_productivity_factors %>%
  full_join(uk_gdp_focus, by = c("time")) %>%
  full_join(uk_productivity_whole, by = c("time")) %>%
  select(c("time", "r_d_exp_measure", "r_d_exp_value","uk_productivity_value", "GDP")) %>%
  filter(r_d_exp_measure == "As a % of GDP") %>%
  distinct() %>%
  ggplot() +
  aes(x = time, y = uk_productivity_value) +
  geom_line(group = 1, colour =   "#91bfdb", size = 2) +
  geom_point(colour =   "#91bfdb", size = 3) +
     labs(x = "\nYear",
         y = "Output per Hour Worked") +
    theme_classic(base_size = 10) +
    theme(plot.title = element_text(hjust = 0.5))
```

```{r}
# graph saved manually
grid.newpage()
grid.draw(rbind(ggplotGrob(r_d_graph_1), ggplotGrob(r_d_graph_2), size = "last"))
```

```{r}
r_d_graph_3 <- uk_productivity_factors %>%
  full_join(uk_gdp_focus, by = c("time")) %>%
  full_join(uk_productivity_whole, by = c("time")) %>%
  select(c("time", "r_d_exp_measure", "r_d_exp_value","uk_productivity_value", "GDP")) %>%
  filter(r_d_exp_measure == "As a % of GDP") %>%
  mutate(r_and_d_spend = (GDP*r_d_exp_value)/(100*1000)) %>%
  distinct() %>%
  ggplot() +
  aes(x = r_and_d_spend, y = uk_productivity_value) +
  geom_point(colour = "#91bfdb", size = 2) + 
  geom_smooth(method = "lm", se = FALSE, colour = "#fc8d59", size = 2) +
     labs(x = "\nR&D Investment (£ billion)",
         y = "Output per Hour Worked",
         title = "R and D Investment vs UK Productivity") +
    theme_classic(base_size = 18) +
    theme(plot.title = element_text(hjust = 0.5))

ggsave("presentation/r_d_investment_vs_productivity_graph.png", plot = r_d_graph_3, width = 11)
```



## Government Investment vs Productivity


### Infrastructure Investment
```{r}
cross_country_infrastructure_investment <- all_international_data %>%
  select(c("time", "location", "labour_productivity_measure", "labour_productivity_value", "total_investment", "population")) %>%
  filter(!is.na(total_investment),
         total_investment > 0, # one point for Norwary was set to zero, unlikely
         labour_productivity_measure == "USD",
         location %in% c("LUX", "BEL", "NOR", "NLD", "USA", "GBR")) %>% # compare UK against countries that are better than use in terms of productivity
  mutate(investment_per_pop = total_investment/population) %>%
  distinct() %>%
  ggplot() +
  aes(x = investment_per_pop, y = labour_productivity_value, colour = location) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE)+
  labs(x = "\nInfrastructure Investment (US Dollars /per person)",
           y = "Labour Productivity (GDP per hour worked)",
           title = "Infrastructure Investment per Population vs Labour Productivity",
       colour = "Country") +
      theme_classic(base_size = 18) +
      theme(plot.title = element_text(hjust = 0.5))

ggsave("presentation/cross_country_infrastructure_investment_graph.png", plot = cross_country_infrastructure_investment, width = 11)
```

### Intangible Assets

```{r}
intangible_assets_graph <- uk_productivity_factors %>%
  full_join(uk_productivity_whole, by = "time") %>%
  select(c("time", "uk_productivity_value","asset_category", "asset_value_billions")) %>%
  filter(asset_category == "Total") %>%
  distinct() %>%
  ggplot() +
  aes(x = asset_value_billions, y = uk_productivity_value) +
  geom_point(colour = "#91bfdb", size = 2) + 
  geom_smooth(method = "lm", se = FALSE, colour = "#fc8d59", size = 2) +
  labs(x = "\nIntangible Asset Investment (£ billions)",
             y = "Output per hour Worked",
             title = "Intangible Asset Investment vs UK Productivity",
         colour = "Country") +
        theme_classic(base_size = 18) +
        theme(plot.title = element_text(hjust = 0.5))

ggsave("presentation/intangible_assets_graph.png", plot = intangible_assets_graph, width = 11)
```

























